An efficient acyclic contact planner for multiped robots

### 2017-01-10 body/conclusion.tex
overview + part 4

>>>>

Again, we assume that, for \gls{cluttered} problems, most of the times \contactreachability\, implies \equilibriumfeasibility.
Under this assumption, our goal is to find a \gls{contact reachable} guide path.

To generate such a path efficiently, ideally we need to only sample \gls{contact reachable} root configurations.
This requires exhibiting a necessary and sufficient condition for \contactreachability. %,that does not need the computation of the whole-body configuration.
By default, verifying \contactreachability\, implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009},
which is too computationally expensive.
To formulate a cheaper condition, we may turn our attention towards either an only-necessary or an only-sufficient condition.

Only-necessary conditions are appealing because they preserve the completeness of the search, while reducing 
the search space: they provide an outer approximation of $C_{Contact}^0$.
On the other hand, only-sufficient conditions provide the guarantee that any configuration that satisfies them is indeed \gls{contact reachable}:
they provide an inner approximation of $C_{Contact}^0$.

In practice, the only-necessary and only-sufficient conditions that we can provide are trivial and give rather inaccurate approximations of $C_{Contact}^0$.
Therefore, we propose a compromise between them: the \textit{reachability condition}, which is computationally efficient
and provides a rather accurate approximation of $C_{Contact}^0$.

### 2017-01-10 body/contact.tex
typos
nicolas comments
overview + part 4

% !TEX root =  ../main.tex
\section{From a guide path to a discrete sequence of contact configurations ($\mathcal{P}_2$)}
\label{sec:contact}
Our planner computes guide paths in \gls{creach0}, an approximation of \gls{ccequil0}.
As an input of this stage, we however assume an \gls{equilibrium feasible} root guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ \gls{ccequil0}.
If this is not the case, our planner will fail rapidly, thus allowing replanning, as discussed in Section~\ref{sec:perf}.
We now consider the second problem of computing a discrete sequence of equilibrium configurations $\mathbf{Q}^{\overline{0}}$ along $\mathbf{q}^0(t)$.

@@

\subsubsection{Maintaining a contact in the sequence.}

Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If possible, a limb in contact at step $i-1$ remains in contact at step $i$. The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. The solver is directly provided by the Hpp software.

@@

%\subsection{Generation of a contact configuration}  
\subsubsection{Creating contacts.}
Contacts are created using three rules:
\begin{enumerate}
\item Only one contact creation can happen between two consecutive steps. This increases the odds that the interpolation between the two steps be feasible;
%\adnote{Before you said that you ensured that transition between two states is feasible...}
\item A contact is validated if and only if the resulting configuration is in static equilibrium; 
\item We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others.
\end{enumerate}

>>>>

% !TEX root =  ../main.tex
\section{From a guide path to a discrete sequence of contact configurations ($\mathcal{P}_2$)}
\label{sec:contact}
Our planner computes guide paths in $C_{Reach}^0$ , an approximation of \gls{$C_{Equil}^0$}.
As an input of this stage, we however assume an \gls{equilibrium feasible} root guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ \gls{$C_{Equil}^0$}.
If this is not the case, our planner will fail rapidly, thus allowing replanning, as discussed in Section~\ref{sec:perf}.
We now consider the second problem of computing a discrete sequence of equilibrium configurations $\mathbf{Q}^{\overline{0}}$ along $\mathbf{q}^0(t)$.

@@

\subsubsection{Maintaining a contact in the sequence:}

Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If possible, a limb in contact at step $i-1$ remains in contact at step $i$. The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. The solver is directly provided by the HPP software.

@@

%\subsection{Generation of a contact configuration}  
\subsubsection{Creating contacts:}
Contacts are created using three rules:
\begin{enumerate}
\item Only one contact creation can happen between two consecutive steps. This increases the odds that the interpolation between the two steps be feasible;
%\adnote{Before you said that you ensured that transition between two states is feasible...}
\item A contact is validated if and only if the resulting configuration is in static equilibrium; 
\item We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others.
\end{enumerate}

### 2017-01-10 body/overview.tex
overview + part 4

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a root guide path with the \textit{reachability condition}, before
extending it into a sequence of configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper.

@@

\subsection{Notation conventions and definitions} \label{notations}

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, that we divide into different parts: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DOFS: for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing $R^0$ is denoted $W^0$ (Figure~\ref{fig:HyQ_roms}).
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

\medskip
\textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains.
%~ Finally, we define $dist(\mathbf{x}, O)$ as the minimal distance between a point $\mathbf{x}$ and the closest surface of the environment.

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: we first plan a root guide path with the \textit{reachability condition}, before
extending it into a sequence of configurations in static equilibrium.

@@

\section{Notation conventions and definitions} \label{notations}

@@

\medskip
\textbf{A robot} is a kinematic chain $R$ composed of: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DOFS: for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$.
$\mathbf{q}$ is decomposed as follows:
\begin{itemize}
	\item $\mathbf{q}^k$ is a configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ is a vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ is the world coordinates vector of $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
\noindent $W^0$ is a volume encompassing $R^0$ (Figure~\ref{fig:HyQ_roms}). \\
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 $W^k$ is the reachable workspace of a limb $R^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

\medskip
\textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains.

\medskip
$C_{Contact}$ is the set of whole body-configurations in contact and collision-free.

\medskip
$C_{Equil} \subset C_{Contact}$ is the set of whole body-configurations in static equilibrium and collision-free.

\medskip
For any set $C_{X}$, we define $C_{X}^0$:
\begin{equation*}
  C_{X}^0 = \left\{ \mathbf{q}^{0},  \exists \mathbf{q}^{\overline{0}}: \mathbf{q}^0  \oplus \mathbf{q}^{\overline{0}} \in C_{X} \right\}
\end{equation*}

%~ Finally, we define $dist(\mathbf{x}, O)$ as the minimal distance between a point $\mathbf{x}$ and the closest surface of the environment.

### 2017-01-10 body/rbprm.tex
overview + part 4

% !TEX root =  ../main.tex
\section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\label{rbprm}

We consider the issue of computing an \gls{equilibrium feasible} guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ \gls{ccequil0} for the root of a multiped robot, connecting user-defined start and goal configurations.

Again, we assume that, for \gls{cluttered} problems, most of the times \contactreachability\, implies \equilibriumfeasibility.
Under this assumption, our goal is to find a \gls{contact reachable} guide path.

To generate such a path efficiently, ideally we need to only sample \gls{contact reachable} root configurations.
This requires exhibiting a necessary and sufficient condition for \contactreachability. %,that does not need the computation of the whole-body configuration.
By default, verifying \contactreachability\, implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009},
which is too computationally expensive.
To formulate a cheaper condition, we may turn our attention towards either an only-necessary or an only-sufficient condition.

Only-necessary conditions are appealing because they preserve the completeness of the search, while reducing 
the search space: they provide an outer approximation of \gls{ccontact0}.
On the other hand, only-sufficient conditions provide the guarantee that any configuration that satisfies them is indeed \gls{contact reachable}:
they provide an inner approximation of \gls{ccontact0}.

In practice, the only-necessary and only-sufficient conditions that we can provide are trivial and give rather inaccurate approximations of \gls{ccontact0}.
Therefore, we propose a compromise between them: the \textit{reachability condition}, which is computationally efficient
and provides a rather accurate approximation of \gls{ccontact0}. % for \gls{cluttered} problems.

%~ As said in the previous section, the goal is to enforce that the guide can be discretized into a sequence of \gls{equilibrium feasible} configurations. We denote by \gls{ccequil} the equilibrium sub-manifold of the robot.

%~ We say that a root configuration $\mathbf{q}^{0}$ is \gls{equilibrium feasible} if it belongs to the set:
%~ \begin{equation}
      %~ C_{Equil}^0= \{ \mathbf{q}^{0}: \exists \mathbf{q}^{\overline{0}}, \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{Equil} \}
    %~ \label{eq:pi}
%~ \end{equation}

%~ The set of all \textit{geometrically feasible} root configurations is denoted by $C_{reach}$.
%~ By extension, a path $\mathbf{q}^0(t)$ is \gls{equilibrium feasible} if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{contact}^0$.

%~ In this Section we characterize \gls{ccontact0} theoretically with a necessary and a sufficient condition for belonging to the set.
%~ We then approximate \gls{ccontact0} as a space $C_{reach}$, where we can efficiently sample configurations.
%~ We finally describe our implementation of an efficient RRT planner in $C_{reach}$.

%~ We want all the configurations of $q^0(t)$ to belong to a set $C_{reach}$ defined as:
%~ 
%~ \begin{eqnarray}
%~ C_{reach}= \{ \mathbf{q}^{0} & : & \exists f, f(\mathbf{q}^{0}) \in C_{contact} \}
%~ \end{eqnarray}

\subsection{Conditions for contact reachability}
%~ By default, \glslink{contact reachable}{\textit{contact reachability}} implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing $\mathbf{q}^{\overline{0}}$ is expensive in terms of computation time. In this section we rather define a necessary condition and a sufficient condition for \glslink{contact reachable}{\textit{contact reachability}} that do not require this explicit computation.

\subsubsection*{contact reachability, a necessary condition:}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is \gls{contact reachable}, then the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion \gls{ccontact0} $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\subsubsection*{contact reachability, a sufficient condition:}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact reachable}{\textit{contact reachability}} can be constructed as a variation of $\mathcal{C}^0_{\textrm{\it Nec}}$, by replacing $W^0$ with a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition:

\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

\subsection{contact reachability: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact reachable}. 
An ideal shape $B$ (with $W^0 \subset B \subset B^{\textrm{\it Suf}}$) that leads to a necessary and sufficient condition may exist---even if it seems intuitively very unlikely in general. 
%~ Its construction is out of the scope of this work.

%~ It might be possible to find an ideal shape $B$ that leads to a necessary and sufficient condition; however, it seems intuitively very unlikely in general. The construction of a shape $W^0 \subset B \subset B^{\textrm{\it Suf}}$ leading to a necessary and sufficient condition (or the proof of its nonexistence) is out of the scope of this work.

However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$. We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient.
%~ The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact reachable} configurations can be found). 
The parametrization of $s$ then defines a trade-off between these two interesting extremes. 
We can choose $s$ by hand, or automatically as explained in Section~\ref{sec:params}.
The chosen value $s=s^*$ defines the \gls{reachc}, therefore we write \gls{creach0} $= C^0_{s^*}$.
%~ We call this condition the \gls{reachc}, which is defined by the chosen value $s=s^*$, and define \gls{creach0} $= C^0_{s^*}$.
%~ In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

%~ As an example, Figure~\ref{fig:HyQ_roms} presents the volumes $W$ computed for the HyQ robot.
In Appendix~\ref{app:rom}, we give a generic method to compute the volumes appearing in the definition of \gls{creach0}, with the example of HRP-2.
%~ Section~\ref{sec:results} shows that in practice, it is easy to adjust $s$ to keep most of the interesting guides without introducing incorrect guides.

\subsection{Computing the guide path in $C_{reach}^0$}
Any sampling-based motion planner can be used to plan a path in \gls{creach0}. 
Indeed, contrary to \gls{ccontact}, \gls{creach0} has a nonzero measure in the configuration space $C$. Therefore a standard uniform sampling approach
can work, in spite of a high rejection rate. 
Thus, the only significant change regarding a classical planner is to replace the collision checking with the \gls{reachc} when verifying
the drawn configurations and associated local paths.

@@

Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso). By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.

>>>>

% !TEX root =  ../main.tex
%~ \section{Computing a guide path in $C^0_{reach}$ ($\mathcal{P}_1$) }
\section{Root path planning in the \textit{contact reachable space}}
\label{rbprm}

In the root path planning phase of our framework, we only consider the root configuration $\mathbf{q}^0$ defined in the previous Section,
as well as the environment $O$.

Given a start and a goal configurations, we aim at computing a guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow$ $\mathbb{R}^r$ verifying:
\begin{equation*} \label{eq:path}
\forall t \in [0,1], \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation*}
This means that any root configuration must be extended into a whole-body, static equilibrium configuration.
$C_{Equil}^0$  can't be described analytically, because the contact manifold has a zero measure.

The strong hypothesis of this work is that for a large variety of locomotion tasks, we can define a space  $C_{Reach}^0 \simeq C_{Contact}^0$, such that 
\begin{equation} \label{eq:creach}
\forall t \in [0,1], \mathbf{q}^0(t) \in C_{Reach}^0 \Rightarrow \mathbf{q}^0(t)  \in C_{Equil}^0
\end{equation}
% for \gls{cluttered} problems.
We call  $C_{Reach}^0$ the \textit{contact reachable workspace}, and detail its construction in the following.
The validity of this hypothesis is discussed in depth in Section~\ref{sec:discussion}.
%~ We describe the construction of $C_{Reach}^0$ in the following.

\subsection{Conditions for contact reachability}
The contact reachable workspace is defined as a compromise between two necessary and a sufficient condition for contact creation.

\textbf{necessary conditions:}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects the reachable workspace $W(\mathbf{q}^{0})$ of the robot (Figure~\ref{fig:contact_gen}--1). Also the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily collision-free.
Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ $C_{Contact}^0$ defined as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion $C_{Contact}^0$ $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

@@

\textbf{a sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces in contact. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

@@

\subsection{The compromising reachability condition}
The ideal shape $B^*, W^0 \subset B* \subset B^{\textrm{\it Suf}}$ that defines a necessary \textbf{and} sufficient condition for contact creation has no analytical definition.
Therefore, we approximate $B^*$ to define the contact reachable space $C_{Reach}^0$.

%~ However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. 
We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
 \label{eq:reachc}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The parametrization of $s$ defines a trade-off:
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$.
 %~ We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient, but less and less necessary.  
Eq.~\ref{eq:reachc} thus defines the \textit{reachability condition}. We fix a value $s^*$ for $s$ and define  $C_{Reach}^0 = C^0_{s^*}$.
The computation of $s^*$ is detailed in Section~\ref{sec:params}. 
In Appendix~\ref{app:rom}, we give a generic method to compute the $W$ volumes appearing in the definition of $C_{Reach}^0$.

\subsection{Computing the guide path in $C_{reach}^0$}
$C_{Reach}^0$ can be sampled efficiently thanks to Eq.~\ref{eq:reachc}, and can thus be used with a standard motion planner.
Thus, the only significant change regarding a classical planner is to replace the collision checking with the \textit{rechability condition} when verifying
the drawn configurations and associated local paths.

@@

Thanks to these modifications, the problem of planning a \gls{contact reachable} path is reduced to a geometric collision-checking problem, of low dimension (6 for HyQ, 8 for HRP-2 that has 2 joints in the torso).
%~ By doing this we can solve the problem with a sampling-based approach, in \gls{interactive} computation times.

### 2017-01-10 body/results.tex
overview + part 4

\subsubsection{Choosing the scaling factor $s$:} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (i.e. the configuration satisfies
the reachability condition, but does not belong to \gls{ccequil0}) and false negatives (i.e. the configuration does not satisfy the 
reachability condition, but belongs to \gls{ccequil0}).

>>>>

\subsubsection{Choosing the scaling factor $s$:} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (i.e. the configuration satisfies
the reachability condition, but does not belong to \gls{$C_{Equil}^0$}) and false negatives (i.e. the configuration does not satisfy the 
reachability condition, but belongs to \gls{$C_{Equil}^0$}).
